Online Social Cohesion Reflects Real-World Group Action in Syria during the Arab Spring

### 2020-06-05 manuscript-chicago.Rmd
Removing more language

The use of similar language can provide insight into *interlocutor cohesion*, or
the degree to which people are talking about similar things. This can---as noted
earlier---include people who are talking about similar things from radically
perspectives, as well as those who are talking about something from the same
perspective. Previous research from political science has investigated these
ideas [@Windsor2020]: The “Occupy Wall Street” movement has been held up as an
example of a movement with low interlocutor and linguistic cohesion because of
its wide range of signs and slogans, while the “I am a man” protest (captured by
Ernest Withers during the 1968 sanitation workers' strike in Memphis, TN) or the
“Black Lives Matter” placards (following the uprising in Ferguson, MO) have high
lexical and interlocutor cohesion [@edgar2018struggle].

Fed by lexical cohesion, interlocutor cohesion ebbs and flows along a timeline
contextualized by variation in real world events. Interlocutor cohesion may
increase preceding an in-person social mobilization event, alerting a repressive
government that collective mobilization is happening. Interlocutor cohesion may
also respond to exogenous events such as spontaneous, unprovoked militarized
conflict. If government forces target a new location using military weapons,
this can precipitate online mobilization and discussion of the event.

Given the importance of lexical and interlocutor cohesion for messaging in
political movements, lexical cohesion may be an important bit of trace data for
researchers interested in leveraging social media data to analyze social
movements. Similar to large-scale studies of sociocultural movements [sometimes
referred to as *culturomics*; e.g., @michel2011quantitative], political
scientists can track the ways in which conversations on social media coalesce or
diffuse around similar content words. In this way, it matters relatively little
whether the interlocutors have precisely the same views: The point is to
understand whether the society-level conversation is converging on similar
language.

@@

This social cohesion can have real and wide-ranging impacts on a political
movement and the people in it. A socially cohesive group is inherently more
threatening to the existing regime because it reflects collective opposition
momentum and the professionalization of potential alternate leadership. Groups
with high social cohesion are more difficult to divide and conquer, and more
cohesive messages may also garner increased international support. Global
insurgency movements exhibit similar patterns in their social structure
[@bohorquez2009common]. Each violent group has a stable state. Exogenous events
can create perturbations that push the group to become either more cohesive or
more fragmented. Rather than frenetic, patchwork messages via Twitter, emergent
or intentional cohesive online messaging---using similar language, paralleling
in-person messaging [@Windsor2020]---are a sign of increased organizational
structure, inciting a government response.

@@

Political and social cohesion in the world can be modeled using event data which
encodes who does what to whom on a given day. This could include events such as
the government imposing a blockade or fighting with artillery or tanks on
a civilian group. CAMEO (Conflict and Mediation Event Observations) encode
a set of activities with numerical values, rescaled to reflect a
conflict-cooperation scale [@schrodt2006twenty; @goldstein1992conflict]. These event
codes include a range of behaviors, from making public statements and appeals,
to demanding, threatening, protesting, and using unconventional mass violence.

### Sociolinguistic Cohesion: Connectedness of Speakers, Ideas, and Events
We tie the constructs of lexical, interlocutor, and social cohesion together in
a theoretical framework called *sociolinguistic cohesion*, as shown in Figure
\@ref(fig:theoretical-framework). Several interconnected mechanisms are in
motion here. Previous values of linguistic cohesion should influence future
values of linguistic cohesion and future actions in the real world space.
Previous events influence future events and future linguistic cohesion in
social media.

@@

Sociolinguistic cohesion is not a static concept. Just as social movements
experience increased participation and setbacks---indicating dynamic
relationships among participants and between participants and the sociopolitical
environment---sociolinguistic cohesion is also dynamic, reflecting the changing
states of the political sphere. Other scholars have noted the diffusion of
messages whereby they reach a critical mass, whereby “behaviors, beliefs, and
preferences” spread within a population [@barash2012critical].  Cohesion
changes over time in response to external events and the internal dynamics of
the social movement. This can occur even if the social interaction occurs entirely 
online [@centola2007complex].

@@

Combining these separate threads, it is reasonable to expect that real-world
actions and online social interactions would be enmeshed, driving and being
driven by one another. Rather than expecting that only real-world social
cohesion would be driven by real-world action, we would expect interaction
between real-world and online dynamics: In this case, exposure to real-world
violence could result in social cohesion in the virtual world. Increased social
cohesion is thus potentially an indicator and a result of increased social
organization and mobilization on the ground. However, this occurs within a
context that is affected by local and international events; both exogenous
events (like intervention by another country or international organization like
the United Nations) and endogenous events (like the recruitment of new members
to the rebel or resistance movement) can affect cohesion.

From a dynamical systems perspective, online social cohesion and in-person
action can be considered two complementary components of a single system,
mutually constrained and constantly interacting and influencing one another
[@mahner1997foundations]. Investigation of the coupling of real-world
mobilization and social cohesion could provide new insight into the power of
social media during mobilization and the ways in which individuals utilize it to
bring about, or respond to, social change. Increasingly, social media platforms
are being utilized during social mobilization [@mcadam2003dynamics;
@koopmans2004movements], but the relationship between virtual and real-world
mobilization is still largely unknown. The present work aims to understand how
physical and virtual movements coexist and interact from a dynamical systems
perspective.

Lexical cohesion, then, is be a useful indicator of social cohesion in
political crisis. Lexical cohesion happens at the word level. Interlocutor
cohesion happens between speakers. Political cohesion happens at the social
level during moments of political crisis, like civil wars and social movements.
Sociolinguistic cohesion is comprised of lexical, interlocutor, and
socio-political events, where the words and their context reveal the disposition
of socially connected linguistic communities. Sociolinguistic cohesion combines
the elements of lexical cohesion in that it captures the degree of similarity
between what is said, and how it is said. It accounts for the cohesion between
multiple speakers in an environment whose language evolves to become more
aligned and similar. Sociolinguistic cohesion also addresses how groups involved
in political rebellion must have internal cohesion to overcome the collective
action problem to mount sustained resistance to state authority.

>>>>

The use of similar language can provide insight into *interlocutor cohesion*, or
the degree to which people are talking about similar things. This can include
people who are talking about similar things from radically perspectives, as well
as those who are talking about something from the same perspective. Previous
research from political science has investigated these ideas [@Windsor2020]: The
“Occupy Wall Street” movement has been held up as an example of a movement with
low interlocutor and linguistic cohesion because of its wide range of signs and
slogans, while the “I am a man” protest (captured by Ernest Withers during the
1968 sanitation workers' strike in Memphis, TN) or the “Black Lives Matter”
placards (following the uprising in Ferguson, MO) have high lexical and
interlocutor cohesion [@edgar2018struggle].

Fed by lexical cohesion, interlocutor cohesion ebbs and flows along a timeline
contextualized by variation in real world events. Interlocutor cohesion may
increase preceding an in-person social mobilization event, alerting a repressive
government that collective mobilization is happening. Interlocutor cohesion may
also respond to exogenous events such as spontaneous, unprovoked militarized
conflict.

Given the importance of lexical and interlocutor cohesion for messaging in
political movements, lexical cohesion may be important trace data for
researchers interested in leveraging social media data to analyze social
movements. Similar to large-scale studies of sociocultural movements [sometimes
referred to as *culturomics*; e.g., @michel2011quantitative], political
scientists can track the ways in which conversations on social media coalesce or
diffuse around similar content words. In this way, it matters relatively little
whether the interlocutors have precisely the same views: The point is to
understand whether the society-level conversation is converging on similar
language.

@@

This social cohesion can have real and wide-ranging impacts on a political
movement and the people in it. A socially cohesive group is inherently more
threatening to the existing regime because it reflects collective opposition
momentum and the professionalization of potential alternate leadership. Global
insurgency movements exhibit similar patterns in their social structure
[@bohorquez2009common]. Rather than frenetic, patchwork messages via Twitter,
emergent or intentional cohesive online messaging---using similar language,
paralleling in-person messaging [@Windsor2020]---are a sign of increased
organizational structure, inciting a government response.

@@

Political and social cohesion in the world can be modeled using event data which
encodes who does what to whom on a given day. CAMEO (Conflict and Mediation
Event Observations) encode a set of activities with numerical values, rescaled
to reflect a conflict-cooperation scale [@schrodt2006twenty;
@goldstein1992conflict]. These event codes include a range of behaviors, from
making public statements and appeals, to demanding, threatening, protesting, and
using unconventional mass violence.

### Sociolinguistic Cohesion: Connectedness of Speakers, Ideas, and Events
We tie the constructs of lexical, interlocutor, and social cohesion together in
a theoretical framework called *sociolinguistic cohesion*, as shown in Figure
\@ref(fig:theoretical-framework). Previous values of linguistic cohesion should influence future values of linguistic cohesion and future actions in the real world space.
Previous events influence future events and future linguistic cohesion in
social media.

@@

Sociolinguistic cohesion is not a static concept. Just as social movements
experience increased participation and setbacks---indicating dynamic
relationships among participants and between participants and the sociopolitical
environment---sociolinguistic cohesion is also dynamic, reflecting the changing
states of the political sphere. Other scholars have noted the diffusion of
messages whereby they reach a critical mass, allowing for “behaviors, beliefs, and
preferences” to spread within a population [@barash2012critical]. Cohesion
changes over time in response to external events and the internal dynamics of
the social movement, even if the social interaction occurs entirely 
online [@centola2007complex].

@@

Combining these separate threads, it is reasonable to expect that real-world
actions and online social interactions would be enmeshed, driving and being
driven by one another. From a dynamical systems perspective, online social
cohesion and in-person action can be considered two complementary components of
a single system, mutually constrained and constantly interacting and influencing
one another [@mahner1997foundations]. Investigation of the coupling of
real-world mobilization and social cohesion could provide new insight into the
power of social media during mobilization and the ways in which individuals
utilize it to bring about, or respond to, social change. Increasingly, social
media platforms are being utilized during social mobilization
[@mcadam2003dynamics;
@koopmans2004movements], but the relationship between virtual and real-world
mobilization is still largely unknown. The present work aims to understand how
physical and virtual movements coexist and interact from a dynamical systems
perspective.

Lexical cohesion, then, is a useful indicator of social cohesion in
political crisis. Lexical cohesion happens at the word level. Interlocutor
cohesion happens between speakers. Political cohesion happens at the social
level during moments of political crisis, like civil wars and social movements.
Sociolinguistic cohesion is comprised of lexical, interlocutor, and
socio-political events, where the words and their context reveal the disposition
of socially connected linguistic communities. Sociolinguistic cohesion combines
the elements of lexical cohesion in that it captures the degree of similarity
between what is said, and how it is said. It accounts for the cohesion between
multiple speakers in an environment whose language evolves to become more
aligned and similar. Sociolinguistic cohesion also addresses how groups involved
in political rebellion must have internal cohesion to overcome the collective
action problem to mount sustained resistance to state authority.
